---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ukmaps <img src="man/figures/logo.svg" align="right" height="139" alt="" />

<!-- badges: start -->
[![BuyMeACoffee](https://raw.githubusercontent.com/pachadotdev/buymeacoffee-badges/main/bmc-yellow.svg)](https://www.buymeacoffee.com/pacha)
<!-- badges: end -->

*If you need an R package or Shiny dashboard for your team, you can email me or inquiry on [Fiverr](https://www.fiverr.com/sellers/msepulveda3).*

Ukmaps provides simplified maps of the United Kingdom administrative and electoral boundaries. Includes maps for
England, Scotland, Wales, and Northern Ireland.

This is a very early version of the package. More features and boundaries will be added soon.

## Installation

You can install the development version of ukmaps like so:

``` r
remotes::install_github("pachadotdev/ukmaps")
```

## Examples

Yes/No map of London administrative areas:

```{r example, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
library(ukmaps)
library(dplyr)
library(ggplot2)

d <- boundaries %>%
  mutate(
    region_name = if_else(is.na(region_name), "Notr Available", region_name),
    is_london = if_else(region_name == "London", "Yes", "No")
  )

pal <- c("#165976", "#d04e66")

ggplot(d) + 
  geom_sf(aes(fill = is_london, geometry = geometry), color = "white") +
  scale_fill_manual(values = pal, name = "Is this London?") +
  labs(title = "Map of England with Administrative Boundaries") +
  theme_minimal(base_size = 13)
```

Which part of London is Barnet?

```{r example3, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
d <- boundaries %>%
  filter(region_name == "London") %>%
  mutate(is_barnet = if_else(lad_name == "Barnet", "Yes", "No"))

pal <- c("#165976", "#d04e66")

ggplot(d) + 
  geom_sf(aes(fill = is_barnet, geometry = geometry), color = "white") +
  scale_fill_manual(values = pal, name = "Is this Golders Green?") +
  labs(title = "Map of Barnet (London) with Electoral Boundaries") +
  theme_minimal(base_size = 13)
```

Which part of London is Golders Green?

```{r example4, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
d <- boundaries %>%
  filter(region_name == "London") %>%
  mutate(
    is_golders_green = if_else(ward_name == "Golders Green", "Yes", "No")
  )

pal <- c("#165976", "#d04e66")

ggplot(d) + 
  geom_sf(aes(fill = is_golders_green, geometry = geometry), color = "white") +
  scale_fill_manual(values = pal, name = "Is this Golders Green?") +
  labs(title = "Map of Barnet (London) with Electoral Boundaries") +
  theme_minimal(base_size = 13)
```

The following maps use functions that aggregate the dataset to keep the package size small.

Country level map of the UK:

```{r example5, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
pal <- c("#165976", "#365158", "#d04e66", "#ffd613")

# country() aggregates the map to country level
ggplot(country()) + 
  geom_sf(aes(fill = country_name, geometry = geometry), color = "white") +
  scale_fill_manual(values = pal, name = "Country") +
  labs(title = "Map of England with Country Boundaries") +
  theme_minimal(base_size = 13)
```

How many R's in each county name?

```{r example6, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
# number of R's in county names
d <- counties() %>%
  mutate(n = stringr::str_count(county_name, "[rR]"))

# region() aggregates the map to country level
ggplot(d) + 
  geom_sf(aes(fill = n, geometry = geometry), color = "white") +
  scale_fill_gradient(low = "#165976", high = "#d04e66", name = "R's",
                      breaks = seq(0, max(d$n), by = 1)) +
  labs(title = "How many R's in each county name?") +
  theme_minimal(base_size = 13)
```

How many R's in each LAD name? Local Authority Districts (LAD) (Local Government District (LGD) in Northern Ireland)

```{r example7, fig.width=15, fig.height=10, warning=FALSE, message=FALSE}
d <- lads() %>%
  mutate(n = stringr::str_count(lad_name, "[rR]"))

ggplot(d) + 
  geom_sf(aes(fill = n, geometry = geometry), color = "white") +
  scale_fill_gradient(low = "#165976", high = "#d04e66", name = "R's",
                      breaks = seq(0, max(d$n), by = 1)) +
  labs(title = "How many R's in each LAD name?") +
  theme_minimal(base_size = 13)
```
